---
title: "Algebra Lesson Title Goes Here"
    
output: 
  rmdformats::readthedown:
    self_contained: true
  html_notebook:
    number_sections: no
    toc: yes
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
---



```{r include=FALSE}

# Check and install missing libraries

list.of.packages <- c("ggplot2","dplyr", "DT", "readxl", "stringr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[
  ,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Remove temporary variables

rm(list.of.packages)
rm(new.packages)

# Manual packages (run once)
# remotes::install_github("juba/rmdformats")
```

```{r include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, echo=FALSE)
```

# Vocabulary

## Origin

The center of the graph. It is _always_ the point (0,0).

<iframe src="https://www.desmos.com/calculator/4y9p90qzc3?embed" width="300px" height="300px" style="border: 1px solid #ccc" frameborder=0></iframe>

## Quadrant

The four quarters (or '_boxes_') that make up a graph. We use roman numerals 
when naming the four quadrants. The first quadrant is at the top right, 
and we continue counter-clockwise. 

**Quarter** I begins at the top right. Any point in this area is said to be in 
_Quadrant I_.

**Quarter II** begins at the top left. Any point in this area is said to be in 
_Quadrant II_.

**Quarter III** is at the bottom left. Any point in this area is said to be in 
_Quadrant III_.

**Quarter IV** is at the bottom right. Any point in this area is said to be in 
_Quadrant IV_.

<iframe src="https://www.desmos.com/calculator/egpevp9ne4?embed" width="300px" height="300px" style="border: 1px solid #ccc" frameborder=0></iframe>

## X-Axis

The horizontal line on the graph, indicating movement from left to right.

<iframe src="https://www.desmos.com/calculator/dznsj8uvho?embed" width="300px" height="300px" style="border: 1px solid #ccc" frameborder=0></iframe>

## X-Coordinate

The first value in a point, indicating movement from left to right starting at the origin. Beginning at (0,0), a **positive** x-coordinate means go **right**, and a **negative** x-coordinate means go **left**.

<iframe src="https://www.desmos.com/calculator/lzgjcvaq8h?embed" width="300px" height="300px" style="border: 1px solid #ccc" frameborder=0></iframe>

In a table, the x-coordinate is in the first column. Try changing the values of the first column below.

<iframe src="https://www.desmos.com/calculator/hhmje8oflr" width="560px" height="315px" style="border: 1px solid #ccc" frameborder=0></iframe>

## Y-Axis

<iframe src="https://www.desmos.com/calculator/r9hjavxkjj?embed" width="300px" height="300px" style="border: 1px solid #ccc" frameborder=0></iframe>

## Y-Coordinate

<iframe src="https://www.desmos.com/calculator/800pyzlilu?embed" width="300px" height="300px" style="border: 1px solid #ccc" frameborder=0></iframe>

<iframe src="https://www.desmos.com/calculator/sinrytdtuo" width="560px" height="315px" style="border: 1px solid #ccc" frameborder=0></iframe>

## Coordinate Plane

The graph itself created by the intersection of the x- and y-axes. It's the te for the same graph as before.

<iframe src="https://www.desmos.com/calculator/eswnmrqoa6?embed" width="500px" height="500px" style="border: 1px solid #ccc" frameborder=0></iframe>

## Coordinate Pair

The set of both numbers in parentheses _(x,y)_ representing a single _point_ on the coordinate plane.



# Graphing on the Coordinate Plane

## Plotting a point

<b>1. Identify the x- and y-coordinates of the coordinate pair.</b>

The first number of a coordinate pair is the _x-coordinate_. If it's **positive**, you will move **right**, and if it's **negative**, you will move **left**. You _must_ begin by moving horizontally when plotting points.

The second number of a coordinate pair is the _y-coordinate_. If it's **positive**, you will move **up**, and if it's **negative**, you will move **down**.

<b>2. From the origin, follow the coordinates and draw a point.</b>

Go to the origin (0,0) and

1. Move left or right (x-coordinate)

2. Move  up or down (y-coordinate)

3. Draw the point

<b>Example</b>

Plot the coordinate pair (5, -5).

x-coordinate: 5 &#x27A1; move 5 to the right

<iframe src="https://www.desmos.com/calculator/mp3l074ndx?embed" width="500px" height="500px" style="border: 1px solid #ccc" frameborder=0></iframe>

y-coordinate: -5 &#x27A1; move 5 down

<iframe src="https://www.desmos.com/calculator/e8wusbexi4?embed" width="500px" height="500px" style="border: 1px solid #ccc" frameborder=0></iframe>

<iframe src="https://www.desmos.com/calculator/kznisrpz9v?embed" width="500px" height="500px" style="border: 1px solid #ccc" frameborder=0></iframe>

## Practice

<iframe src="https://www.desmos.com/calculator/xghtgwv1au" width="560px" height="315px" style="border: 1px solid #ccc" frameborder=0></iframe>

<details><summary>Show Answer</summary>


<iframe src="https://www.desmos.com/calculator/z0swxadr3v" width="560px" height="315px" style="border: 1px solid #ccc" frameborder=0></iframe>

</details>

<!--

## Power Point Slides

<iframe src="https://docs.google.com/presentation/d/e/2PACX-1vQMBdagNW9dyhEpPG1gSz9f2dPlpFlnGQvd5kYYFru4Ggc25kzdcXQRFegT9iQiJl3VbsUEvkuzIGoY/embed?start=false&loop=false&delayms=3000" frameborder="0" width="560" height="315" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>


## Youtube Video

 <iframe src="https://www.youtube.com/embed/MfNa7J6Zfcs"
  frameborder="0" width="560" height="315" allow="accelerometer; autoplay;
  encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>




<details><summary>Hidden text</summary>no longer hidden</details>

 -->

