---
title: "Laws of Exponents"
    
output: 
  rmdformats::readthedown:
    self_contained: true
    toc_depth: 2
    number_sections: false
    smooth_scroll: yes
  html_notebook:
    number_sections: true
    toc: yes
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

<style>

#content{
    background:#fcfcfc;
    height:100%;
    margin-left:150px;
    /* margin:auto; */
    max-width:1050px;
    min-height:100%;
}

#sidebar {
    color: /*#D9AB16;*/ #B30000;
    background: /*#113458;*/ #B30000;
    width: 150px;
}

#sidebar h2 {
    color: #FFFFFF;
    background-color: #153881;
}

#sidebar a {
    color: /*#3F4548;*/#FFFFFF;
    background-color: #153B81;
}

#sidebar a:hover {
    background-color: /*#3F4548; #153B81;*/ #FFFFFF;
    color: #000000;
    font-weight: bold;
}

#postamble{
    color:#fcfcfc;
    background:#153881;
    border-top:solid 10px #153881;
    font-family:"Lato","proxima-nova","Helvetica Neue",Arial,sans-serif;
    font-size: 90%;
    z-index:400;
    padding:12px;
}

div.highlight_blue { 
    background-color:#e6f0ff;
    border-radius: 5px; 
    padding: 5px;
}

div.highlight_red { 
    background-color:#ffe6ea;
    border-radius: 5px;
    padding: 5px;
}
    
#main .nav-pills {
    margin-top: 20px;
    margin-bottom: 10px;
}

#main .nav-pills > li > a {
    background-color: #EEEEEE;
    background-image: none;
    background-position: 0 0;
    color: #153B81;
}

#main .nav-pills > li > a:hover {
  color: #FFFFFF !important;
  background-color: #153B81;
}

#main .nav-pills > li.active > a,
#main .nav-pills > li.active > a:hover,
#main .nav-pills > li.active > a:focus {
  color: #FFFFFF !important;
  background-color: #153B81;
}

</style>

```{r include=FALSE}

# Check and install missing libraries

list.of.packages <- c("ggplot2","dplyr", "DT", "readxl", "stringr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[
  ,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Remove temporary variables

rm(list.of.packages)
rm(new.packages)

# Manual packages (run once)
# remotes::install_github("juba/rmdformats")
```

```{r include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, echo=FALSE)
```



# Negative Exponents

## Notes

<div class='highlight_blue'>
There are two ways we can think about negative exponents:

1. The base is on the **_wrong side_** of the fraction line.

2. You need to **_divide_** by this number that many times, not multiply.
</div>

If you want to remove the negative sign, flip the base and its exponent to the other side
of the fraction. The negative in the exponent is removed _after_ the flip.

### Example:

$$\begin{align}\frac{3}{2}^{-4}=\frac{1}{2\cdot3^4} && \text{Base 3 flips, power -4 becomes 4.}\end{align}$$

If the negative exponent has a base with parentheses, both sides of the fraction flip 
with **_nothing inside the fraction changing except where it is._**

### Examples:

$$\begin{align}
(\frac{3}{2})^{-4}&=(\frac{2}{3})^{4} && \text{2 and 3 flip.}\\
  (\frac{-2}{-5})^{-4}&=(\frac{-5}{-2})^{4} && \text{-2 and -5 flip.}\\
  (\frac{(3)^{-1}}{(2)^{-3}})^{-4}&=(\frac{2^{-3}}{3^{-1}})^{4} && \text{Keep negative exponents inside.}
\end{align}$$


If you have multiple terms being multipled together, individual bases will move.

Example:

$$\begin{align}\frac{-1^3\cdot3^{-2}}{(-4)^{-2}\cdot7^4}=\frac{-1^3\cdot(-4)^2}{3^2\cdot7^4} 
  && \text{Bases 3 and -4 flip.}\end{align}$$

Below is a video if you would like more review. You can find practice problems on [Khan Academy](https://www.khanacademy.org/math/pre-algebra/pre-algebra-exponents-radicals/pre-algebra-negative-exponents/e/exponents_2) and [IXL](https://www.ixl.com/math/algebra-1/negative-exponents) (click the hyperlinks). You should sign in using your Argosy email before starting.


<iframe width="100%" height="500" src="https://www.youtube.com/embed/JnpqlXN9Whw" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<hr>

# Practice {.tabset .tabset-pills .tabset-fade}

<div class='highlight_red'>
<strong>Directions:</strong>

1. Click _Continue without signing in_ below.

2. Type your first name and last initial.

3. Begin the activity.

</div>

## Mild

<iframe src='https://student.desmos.com/join/m5962r' width='100%' height='615' 
  allowfullscreen='true' mozallowfullscreen='true' webkitallowfullscreen='true'/>

<details><summary>Show Explanation</summary>
Testing
</details>


## Medium

<iframe src='https://student.desmos.com/join/pxray9' width='100%' height='615' 
  allowfullscreen='true' mozallowfullscreen='true' webkitallowfullscreen='true'/>

## Spicy
<!--

## Power Point Slides

<iframe src="https://docs.google.com/presentation/d/e/2PACX-1vQMBdagNW9dyhEpPG1gSz9f2dPlpFlnGQvd5kYYFru4Ggc25kzdcXQRFegT9iQiJl3VbsUEvkuzIGoY/embed?start=false&loop=false&delayms=3000" frameborder="0" width="560" height="315" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>


## Youtube Video

 <iframe src="https://www.youtube.com/embed/MfNa7J6Zfcs"
  frameborder="0" width="560" height="315" allow="accelerometer; autoplay;
  encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>




<details><summary>Hidden text</summary>no longer hidden</details>

 -->

