---
title: "R Notebook</br>State Reporting Q3"
author: "<em>Gerardo Pelayo García</em><br><br>`r format(Sys.time(), '%d %B, %Y')`"
    
output: 
  rmdformats::readthedown:
    self_contained: true
  html_notebook:
    number_sections: no
    toc: yes
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

# Set-Up {.tabset .tabset-fade .tabset-pills}

```{r}

# Check and install missing libraries

list.of.packages <- c("ggplot2","dplyr", "DT", "readxl", "stringr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[
  ,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Remove temporary variables

rm(list.of.packages)
rm(new.packages)

# Manual packages (run once)
# remotes::install_github("juba/rmdformats")
```

```{r}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE)
```

# Functions {.tabset .tabset-fade .tabset-pills}

```{r}
# functions
numbers_only <- function(x) !grepl('\\D', x)

letters_only <- function(x) !grepl('[A-Za-z]', x)

numextract <- function(string){
  require(stringr)
  str_extract(string, "\\-*\\d+\\.*\\d*")
}

total_jdf <- function(vector, with_blanks=TRUE){
  sum <- stack(setNames(strsplit(vector, ','), df$program))
  
  sum <- rename(sum, 
    count=values,
    program=ind)

    sum <- sum[!is.na(sum$count),]

    if(with_blanks){
    temp1 <- sum[grepl('jdf', tolower(sum$count)),]
    temp2 <- sum[numbers_only(sum$count),]
    sum <- rbind(temp1, temp2)
    } else{
    sum <- sum[grepl('jdf', tolower(sum$count)),]
    }

    sum$count <- numextract(sum$count)

    sum <- sum(as.numeric(sum$count))
    
    return(cat(sum))
}
```

# Results (with unidentified) {.tabset .tabset-fade .tabset-pills}

Outcomes that include counts with unspecified funding.
```{r}
# Load data
library(readxl)
library(dplyr)

df <- read_excel(
  'C:/Users/Shadow/Documents/Data/State Reporting/fy20_q3_jdf.xlsx')

df <- rename(df,
  program=`JDF Quarter 3 FY2019-2020`,
  youth_work=`Youth Work`,
  entered_jobs=`Entered Jobs`,
  retained_jobs=`Retained Jobs`,
  industry_cred=`Industry Cred.`,
  northstar_cred=`NorthStar Cred.`,
  entered_p_s=`Entered P/S`,
  entered_training=`Entered Training`,
  partners=`Business Partners`)

df <- data.frame(df)

# gather data across outcomes for Q3

# Sums WITH unidentified funding totals
# Youth work
total_jdf(df$youth_work)
# Entered jobs
total_jdf(df$entered_jobs)
# Retained jobs
total_jdf(df$retained_jobs)
# Industry credentials 
total_jdf(df$industry_cred)
# Northstar credentials
total_jdf(df$northstar_cred)
# Entered post-secondary 
total_jdf(df$entered_p_s)
# Entered training
total_jdf(df$entered_training)
```

# Results (without unidentified) {.tabset .tabset-fade .tabset-pills}

Outcomes that do _not_ include counts with unspecified funding.

```{r}
# Sums WITHOUT unidentified funding totals
# Youth work
total_jdf(df$youth_work, with_blanks = FALSE)
# Entered jobs
total_jdf(df$entered_jobs, with_blanks = FALSE)
# Retained jobs
total_jdf(df$retained_jobs, with_blanks = FALSE)
# Industry credentials 
total_jdf(df$industry_cred, with_blanks = FALSE)
# Northstar credentials
total_jdf(df$northstar_cred, with_blanks = FALSE)
# Entered post-secondary 
total_jdf(df$entered_p_s, with_blanks = FALSE)
# Entered training
total_jdf(df$entered_training, with_blanks = FALSE)
```

# Results (CALIS) 
 
Outcomes whose data comes from CALIS, not the Excel spreadsheet

```{r}
# Load data
library(readxl)
library(dplyr)


# Data that ignores min. attendance of 12 hours to be included in data
df_calis_all <- read_excel(
  'C:/Users/Shadow/Documents/Data/State Reporting/q3_fy20_ignore_rule.xlsx', 3)

# Outcome: recruitment
df_calis_all <- rename(df_calis_all, 
  id=`Student Id`)
df_calis_all <- subset(df_calis_all, select=c(id))
df_calis_all <- distinct(df_calis_all)

# Total recruited
cat(nrow(df_calis_all))
```

```{r}
# Load data
library(readxl)
library(dplyr)


# Data that does NOT ignore min. attendance of 12 hours to be included in data
df_calis <- read_excel(
  'C:/Users/Shadow/Documents/Data/State Reporting/q3_fy20.xlsx', 3)

# Outcome: recruitment
temp <- rename(df_calis, 
  id=`Student Id`)
temp <- subset(temp, select=c(id))
temp <- distinct(temp)

# Total enrolled
cat(nrow(temp))

# Outcome: MSG (measurable skill gain/high school equivalent (HSE))
temp <- rename(df_calis, 
  id=`Student Id`,
  msg=`First PoP Gain Type`)
temp <- subset(temp, select=c(id, msg))

# Total with MSG
cat(nrow(temp[!is.na(temp$msg),]))

# Outcome: high school credentials
cat(nrow(temp[grepl('HSE', temp$msg),]))

# Outcome: gained an EFL 
cat(nrow(temp[grepl('EFL', temp$msg),]))

# Outcome: work experiences???
```
# Verify CALIS data

Parameters specified within CALIS for each data source (page 1 of Excel sheet)

```{r}
# Load data
library(readxl)
library(dplyr)
library(DT)

# Data: does ignore min. attendance rule
df_calis_all <- read_excel(
  'C:/Users/Shadow/Documents/Data/State Reporting/q3_fy20_ignore_rule.xlsx', 1,
  col_names = FALSE)
datatable(df_calis_all)

# Data: does NOT ignore min. attendance rule
df_calis <- read_excel(
  'C:/Users/Shadow/Documents/Data/State Reporting/q3_fy20.xlsx', 1,
  col_names = FALSE)

datatable(df_calis)

```



 

