---
title: "Solving Equations by Combining Like Terms"
    
output: 
  rmdformats::readthedown:
    self_contained: true
  html_notebook:
    number_sections: no
    toc: yes
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
---



```{r include=FALSE}

# Check and install missing libraries

list.of.packages <- c("ggplot2","dplyr", "DT", "readxl", "stringr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[
  ,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Remove temporary variables

rm(list.of.packages)
rm(new.packages)

# Manual packages (run once)
# remotes::install_github("juba/rmdformats")
```

```{r include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, echo=FALSE)
```
<!--
# Vocabulary

## Like Terms

## Addition Property of Equality

## Division Property of Equality

## Inverse Operations

## Multiplication Property of Equality

## Subtraction Property of Equality
-->


# Examples of Math Properties

<hr>

## Addition Property of Equality

_If two sides of an equation are equal, they will remain equal if we add the same number to each side._

<b>General Example</b>

Let _x_, _y_, and _z_ be any numbers.

$\begin{align}x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  x + z &= y + z && \text{Adding } z \text{ results in 2 equal sides.}
\end{align}$

<b>Specific Example</b>

We can use specific numbers to make this more intuitive, substituting _x_, _y_, and _z_ for any values. 

Let _x = 2_, _y = 2_, and _z = 5_,

$\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  2 &= 2 && \text{Substitute } x \text{ and } y \text{ with their values (2).}\\
  2 + z &= 2+ z && \text{Add } z \text{ to both sides.}\\
  2 + 5 &= 2 + 5 && \text{Substitute } z\text{ for its value (5).}\\
  7 &= 7 && \text{Each side remains equal.}
\end{align}$

If you would like a more thorough explanation, here is a Khan Academy video introducing the property. (You should stop the video by 4:30.)

<iframe width="560" height="315" src="https://www.youtube.com/embed/sYyiinqF8Q4" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<hr>

## Subtraction Property of Equality

_If two sides of an equation are equal, they will remain equal if we subtract the same number from each side._

<b>General Example</b>

Let _x_, _y_, and _z_ be any numbers.

$\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  x - z &= y - z && \text{Subtracting } z \text{ results in 2 equal sides.}
\end{align}$

<b>Specific Example</b>

We can use specific numbers to make this more intuitive, substituting _x_, _y_, and _z_ for any values. 

Let _x = 8_, _y = 8_, and _z = 3_,

$\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  8 &= 8 && \text{Substitute } x \text{ and } y \text{ with their values (8).}\\
  8 - z &= 8 - z && \text{Subtract } z \text{ from both sides.}\\
  8 - 3 &= 8 - 3 && \text{Substitute } z\text{ for its value (3).}\\
  5 &= 5 && \text{Each side remains equal.}
\end{align}$

If we recognize that subtracting is the same as adding a _negative_ number:

$\begin{align}
  8 - 3 &= 8 + (-3)
\end{align}$

Then it is clear that the _Addition_ and _Subtraction Properties of Equality_ are essentially the same.

<hr>

## Multiplication Property of Equality

_If two sides of an equation are equal, they will remain equal if we multiply the same number to each side._

<b>General Example</b>

Let _x_, _y_, and _z_ be any numbers.

$\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  x \cdot z &= y \cdot z && \text{Multiplying } z \text{ results in 2 equal sides.}
\end{align}$

<b>Specific Example</b>

We can use specific numbers to make this more intuitive, substituting _x_, _y_, and _z_ for any values. 

Let _x = 2_, _y = 2_, and _z = 5_,

$\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  2 &= 2 && \text{Substitute } x \text{ and } y \text{ with their values (2).}\\
  2 \cdot z &= 2 \cdot z && \text{Multiply } z \text{ to both sides.}\\
  2 \cdot 5 &= 2 \cdot 5 && \text{Substitute } z\text{ for its value (5).}\\
  10 &= 10 && \text{Each side remains equal.}
\end{align}$

If you would like a more thorough explanation, here is a Khan Academy video introducing the property. (By 2:00 he starts using equations as examples.)

<iframe width="560" height="315" src="https://www.youtube.com/embed/LhcYPnfyQtw" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<hr>

## Division Property of Equality

_If two sides of an equation are equal, they will remain equal if we divide the same number from each side._

<b>General Example</b>

Let _x_, _y_, and _z_ be any numbers.

$\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  \frac{x}{z} &= \frac{y}{z} && \text{Dividing } z \text{ results in 2 equal sides.}
\end{align}$

<b>Specific Example</b>

We can use specific numbers to make this more intuitive, substituting _x_, _y_, and _z_ for any values. 

Let _x = 12_, _y = 12_, and _z = 4_,

$\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  12 &= 12 && \text{Substitute } x \text{ and } y \text{ with their values (12).}\\
  \frac{12}{z} &= \frac{12}{z} && \text{Divide } z \text{ from both sides.}\\
  \frac{12}{4} &= \frac{12}{4} && \text{Substitute } z\text{ for its value (4).}\\
  3 &= 3 && \text{Each side remains equal.}
\end{align}$

If we recognize that division is the same as multiplying by the reciprocal of a number:

$\begin{align}
  \frac{12}{4} &= 12 \cdot \frac{1}{4}
\end{align}$

Then it becomes clear that the _Multiplication_ and _Division Properties of Equality_ are essentially the same.

<hr>

## Practice

Identify which property is being used for each equation.

$\begin{align}
  x + 1 &= 9\\
  -1 & -1\\
  x &= 8
\end{align}$

<details><summary>Show Answer</summary>Subtraction Property of Equality</details>

$\begin{align}
  x - 5 &= 5\\
  +5 & +5\\
  x &= 10
\end{align}$

<details><summary>Show Answer</summary>Addition Property of Equality</details>

$\begin{align}
  \frac{3x}{3} &= \frac{18}{3}\\
  x &= 6
\end{align}$

<details><summary>Show Answer</summary>Division Property of Equality</details>

$\begin{align}
  \frac{x}{5} \cdot 5 &= 4 \cdot5\\
  x &= 20
\end{align}$

<details><summary>Show Answer</summary>Multiplication Property of Equality</details>

<hr>

<br>

# Combining Like Terms

<hr>

## Like Terms on the Same Side

If two or more terms **on the same side of the equation** and 1) share the <u>same variable</u> or 2) are both <u>constants</u>. We are _adding_ theese terms while paying attention to the sign (_+_ or _-_) before each term being combined.


<b>Example with Shared Variables and Constants</b>

$\begin{align}
  1x + 2x + 3x + 4 + 5 &= 6\\
  \color{blue}{1x+2x+3x}+\color{red}{4+5} &= 6 && \text{Identify like terms.}\\
  \color{blue}{6x}+\color{red}{4+5} &= 6 && \text{Combine like terms (same variable).}\\
  \color{blue}{6x}+\color{red}{9} &= 6 && \text{Combine like terms (constants).}
\end{align}$

<b>Example with Different Variables</b>

$\begin{align}
  2a - 4a + 5 &= -3b+b\\
  \color{blue}{2a - 4a} + 5 &= \color{red}{-3b - b} && \text{Identify like terms.}\\
  \color{blue}{-2a} + 5&= \color{red}{-3b - b} && \text{Combine like terms (} a \text{).}\\
  \color{blue}{-2a} + 5 &= \color{red}{-4b} && \text{Combine like terms(} b \text{).}
\end{align}$


Here are examples of equations being solved by combining like terms on the same side.

<iframe width="560" height="315" src="https://www.youtube.com/embed/wskwG4PSntU" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<hr>

## Practice

Simplify each equation by combining like terms on the same side of each equation. Do not solve for the variable(s).

$2x + 3x = 8$

<details><summary>Show Answer</summary>
$\begin{align}
  2x + 3x &= 8 && \text{Combine x terms.}\\
  5x &= 8
\end{align}$
</details>

<br>

$3x + 6x - 3 + 2x = 5y - 4y + 2$

<details><summary>Show Answer</summary>
$\begin{align}
  3x + 6x - 3 + 2x &= 5y - 4y + 2\\
  11x - 3 &= 5y - 4y + 2 && \text{Combine x terms.}\\
  11x - 3 &= y + 2 && \text{Combine y terms.}
\end{align}$
</details>

<br>

$-2a + 4a - b + 7c + 2c = -3 + 4$


<details><summary>Show Answer</summary>
$\begin{align}
  -2a + 4a - b + 7c + 2c &= -3 + 4\\
  2a - b + 7c + 2c &= -3 + 4 && \text{Combine a terms.}\\
  2a - b + 9c &= -3 + 4 && \text{Combine c terms.}\\
  2a - b + 9c &= 1 && \text{Combine constants.}
\end{align}$

</details>

<br>


If you would like additional practice combining like terms, [you can go here](https://www.khanacademy.org/math/algebra/x2f8bb11595b61c86:foundation-algebra/x2f8bb11595b61c86:combine-like-terms/e/combining_like_terms_1).
<hr>

## Like Terms on Opposite Sides

If like terms are on **opposite** sides of an  equation, we use _inverse_ operations to eliminate one.

While it does not matter _which_ term you eliminate, isolating the variable will save us a step. Every example will focus on isolating the variable, but your answers are not necessarily wrong.

<b>Example</b>

$\begin{align}
  3x &= -2 + 2x\\
  \color{blue}{3x} &= -2 \color{blue}{+ 2x} && \text{Identify the like terms.}\\
  \color{red}{-2x} & \text{ }\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }
    \text{ } \color{red}{-2x} && \text{Eliminate +2x by subtracting 2x}\\
  x &= -2
\end{align}$

If a term does not have a sign before it, the term is _positive_. We treat positive terms as addition, and it can be helpful to insert 'plus' signs (_+_) before positive terms to make it easier to see their signs.

$\begin{align}
  4x &= 2x - 5 && \text{4x, 2x are positive with no signs.}\\
  \color{red}{+4x} &= \color{red}{+2x} - 5 && \text{Insert a + sign before 4x, 2x.}\\
  \color{blue}{+4x} &= \color{blue}{+2x} - 5 && \text{Identify like terms.}\\
  -2x &= -2x && \text{Eliminate +2x by subtracting 2x}\\
  2x &= -5
\end{align}$

Finally, we should **always combine like terms on the same side <u>first</u>** whenever possible.


<b>Example</b>

$\begin{align}
  5a + 3a &= 2a + 12\\
  \color{blue}{5a + 3a} &= 2a + 12 && \text{Identify like terms on the same side.}\\
  \color{blue}{8a} &= 2a + 12 && \text{Combine like terms.}\\
  \color{red}{5a} &= \color{red}{2a} + 12 && \text{Identify like terms on opposite sides.}\\
  -2a & -2a && \text{Eliminate positive 2a by subtracting 2a.}\\
  3a &= 12
\end{align}$


Here's an example of an equation being solved by combining like terms on opposite sides.

<iframe width="560" height="315" src="https://www.youtube.com/embed/9teKXGoWIQM" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<hr>

## Practice

Combine like terms on opposite sides of each equation. Your answer may be different and still be correct, but prioritize isolating a variable in your step.

$5a - 3b = 4a + 2$

<details><summary>Show Answer</summary>
$\begin{align}
  \color{blue}{5a} - 3b &= \color{blue}{4a} + 2 \\
  -4a\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }
    & -4a && \text{Eliminate positive 4a by subtracting 4a.}\\
  a - 3b &= 2
\end{align}$
</details>

<br>

$-3x + 2y = 5x + 3$

<details><summary>Show Answer</summary>
$\begin{align}
  \color{blue}{-3x} + 2y &= \color{blue}{5x} + 3 \\
  +3x\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }\text{ }
    & +3x && \text{Eliminate -3x by adding 3x.}\\
  2y  &= 8x + 3
\end{align}$
</details>

<br>

<!--

## Power Point Slides

<iframe src="https://docs.google.com/presentation/d/e/2PACX-1vQMBdagNW9dyhEpPG1gSz9f2dPlpFlnGQvd5kYYFru4Ggc25kzdcXQRFegT9iQiJl3VbsUEvkuzIGoY/embed?start=false&loop=false&delayms=3000" frameborder="0" width="560" height="315" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>


## Youtube Video

 <iframe src="https://www.youtube.com/embed/MfNa7J6Zfcs"
  frameborder="0" width="560" height="315" allow="accelerometer; autoplay;
  encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>




<details><summary>Hidden text</summary>no longer hidden</details>

 -->

