---
title: "Solving Equations by Combining Like Terms"
    
output: 
  rmdformats::readthedown:
    self_contained: true
  html_notebook:
    number_sections: no
    toc: yes
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
---



```{r include=FALSE}

# Check and install missing libraries

list.of.packages <- c("ggplot2","dplyr", "DT", "readxl", "stringr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[
  ,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Remove temporary variables

rm(list.of.packages)
rm(new.packages)

# Manual packages (run once)
# remotes::install_github("juba/rmdformats")
```

```{r include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, echo=FALSE)
```
<!--
# Vocabulary

## Like Terms

## Addition Property of Equality

## Division Property of Equality

## Inverse Operations

## Multiplication Property of Equality

## Subtraction Property of Equality
-->


# Examples of Math Properties

## Addition Property of Equality

_If two sides of an equation are equal, they will remain equal if we add the same number to each side._

<b>General Example</b>

Let _x_, _y_, and _z_ be any numbers.

$$
\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  x + z &= y + z && \text{Adding } z \text{ results in 2 equal sides.}
\end{align}
$$

<b>Specific Example</b>

We can use specific numbers to make this more intuitive, substituting _x_, _y_, and _z_ for any values. 

Let _x = 2_, _y = 2_, and _z = 5_,

$$
\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  2 &= 2 && \text{Substitute } x \text{ and } y \text{ with their values (2).}\\
  2 + z &= 2+ z && \text{Add } z \text{ to both sides.}\\
  2 + 5 &= 2 + 5 && \text{Substitute } z\text{ for its value (5).}\\
  7 &= 7 && \text{Each side remains equal.}
\end{align}
$$

If you would like a more thorough explanation, here is a Khan Academy video introducing the property. (You should stop the video by 4:30.)

<iframe width="560" height="315" src="https://www.youtube.com/embed/sYyiinqF8Q4" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<hr>

## Subtraction Property of Equality

_If two sides of an equation are equal, they will remain equal if we subtract the same number from each side._

<b>General Example</b>

Let _x_, _y_, and _z_ be any numbers.

$$
\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  x - z &= y - z && \text{Subtracting } z \text{ results in 2 equal sides.}
\end{align}
$$

<b>Specific Example</b>

We can use specific numbers to make this more intuitive, substituting _x_, _y_, and _z_ for any values. 

Let _x = 8_, _y = 8_, and _z = 3_,

$$
\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  8 &= 8 && \text{Substitute } x \text{ and } y \text{ with their values (8).}\\
  8 - z &= 8 - z && \text{Subtract } z \text{ from both sides.}\\
  8 - 3 &= 8 - 3 && \text{Substitute } z\text{ for its value (3).}\\
  5 &= 5 && \text{Each side remains equal.}
\end{align}
$$

If we recognize that subtracting is the same as adding a _negative_ number:

$$
\begin{align}
  8 - 3 &= 8 + (-3)
\end{align}
$$

Then it is clear that the _Addition_ and _Subtraction Properties of Equality_ are essentially the same.

<hr>

## Multiplication Property of Equality

_If two sides of an equation are equal, they will remain equal if we multiply the same number to each side._

<b>General Example</b>

Let _x_, _y_, and _z_ be any numbers.

$$
\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  x \cdot z &= y \cdot z && \text{Multiplying } z \text{ results in 2 equal sides.}
\end{align}
$$

<b>Specific Example</b>

We can use specific numbers to make this more intuitive, substituting _x_, _y_, and _z_ for any values. 

Let _x = 2_, _y = 2_, and _z = 5_,

$$
\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  2 &= 2 && \text{Substitute } x \text{ and } y \text{ with their values (2).}\\
  2 \cdot z &= 2 \cdot z && \text{Multiply } z \text{ to both sides.}\\
  2 \cdot 5 &= 2 \cdot 5 && \text{Substitute } z\text{ for its value (5).}\\
  10 &= 10 && \text{Each side remains equal.}
\end{align}
$$

If you would like a more thorough explanation, here is a Khan Academy video introducing the property. (By 2:00 he starts using equations as examples.)

<iframe width="560" height="315" src="https://www.youtube.com/embed/LhcYPnfyQtw" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<hr>

## Division Property of Equality

_If two sides of an equation are equal, they will remain equal if we divide the same number from each side._

<b>General Example</b>

Let _x_, _y_, and _z_ be any numbers.

$$
\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  \frac{x}{z} &= \frac{y}{z} && \text{Dividing } z \text{ results in 2 equal sides.}
\end{align}
$$

<b>Specific Example</b>

We can use specific numbers to make this more intuitive, substituting _x_, _y_, and _z_ for any values. 

Let _x = 12_, _y = 12_, and _z = 4_,

$$
\begin{align}
  x &= y && \text{Numbers } x \text{ and } y \text{ are equal.}\\
  12 &= 12 && \text{Substitute } x \text{ and } y \text{ with their values (12).}\\
  \frac{12}{z} &= \frac{12}{z} && \text{Divide } z \text{ from both sides.}\\
  \frac{12}{4} &= \frac{12}{4} && \text{Substitute } z\text{ for its value (4).}\\
  3 &= 3 && \text{Each side remains equal.}
\end{align}
$$

If we recognize that division is the same as multiplying by the reciprocal of a number:

$$
\begin{align}
  \frac{12}{4} &= 12 \cdot \frac{1}{4}
\end{align}
$$

Then it becomes clear that the _Multiplication_ and _Division Properties of Equality_ are essentially the same.

<hr>

## Practice

Identify which property is being used for each equation.

$$
\begin{align}
  x + 1 &= 9\\
  -1 & -1\\
  x &= 8
\end{align}
$$

<details><summary>Show Answer</summary>Subtraction Property of Equality</details>

$$
\begin{align}
  x - 5 &= 5\\
  +5 & +5\\
  x &= 10
\end{align}
$$

<details><summary>Show Answer</summary>Addition Property of Equality</details>

$$
\begin{align}
  \frac{3x}{3} &= \frac{18}{3}\\
  x &= 6
\end{align}
$$

<details><summary>Show Answer</summary>Division Property of Equality</details>

$$
\begin{align}
  \frac{x}{5} \cdot 5 &= 4 \cdot5\\
  x &= 20
\end{align}
$$

<details><summary>Show Answer</summary>Multiplication Property of Equality</details>

<hr>



# Combining Like Terms to Solve Equations

$$
\begin{align}
  3+x &=4 && \text{(Solve for} x \text{.)}\\
  x &=4-3 && \text{(Subtract 3 from both sides.)}\\
  x &=1   && \text{(Yielding the solution.)}
\end{align}
$$
<!--

## Power Point Slides

<iframe src="https://docs.google.com/presentation/d/e/2PACX-1vQMBdagNW9dyhEpPG1gSz9f2dPlpFlnGQvd5kYYFru4Ggc25kzdcXQRFegT9iQiJl3VbsUEvkuzIGoY/embed?start=false&loop=false&delayms=3000" frameborder="0" width="560" height="315" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>


## Youtube Video

 <iframe src="https://www.youtube.com/embed/MfNa7J6Zfcs"
  frameborder="0" width="560" height="315" allow="accelerometer; autoplay;
  encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>




<details><summary>Hidden text</summary>no longer hidden</details>

 -->

