---
title: "Zero Exponents"
    
output: 
  rmdformats::readthedown:
    toc_depth: 1
    number_sections: false
    smooth_scroll: yes
    self-contained: true
  html_notebook:
    number_sections: true
    toc: yes
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

<style>

#content{
    background:#fcfcfc;
    height:100%;
    margin-left:250px;
    /* margin:auto; */
    max-width:1050px;
    min-height:100%;
}

#sidebar {
    color: /*#D9AB16;*/ #B30000;
    background: #153881;
    width: 250px;
}

#sidebar h2 {
    color: #FFFFFF;
    background-color: #153881;
}

#sidebar a {
    color: /*#3F4548;*/#FFFFFF;
    background-color: #153B81;
}

#sidebar a:hover {
    background-color:#FFFFFF;
    color: #000000;
    font-weight: bold;
}

#postamble{
    color:#fcfcfc;
    background:#153881;
    border-top:solid 10px #153881;
    font-family:"Lato","proxima-nova","Helvetica Neue",Arial,sans-serif;
    font-size: 90%;
    z-index:400;
    padding:12px;
}

div.highlight_blue { 
    background-color:#e6f0ff;
    border-radius: 5px; 
    padding: 5px;
}

div.highlight_red { 
    background-color:#ffe6ea;
    border-radius: 5px;
    padding: 5px;
}
    
#main .nav-pills {
    margin-top: 20px;
    margin-bottom: 10px;
}

#main .nav-pills > li > a {
    background-color: #EEEEEE;
    background-image: none;
    background-position: 0 0;
    color: #9F2042;
}

#main .nav-pills > li > a:hover {
  color: #FFFFFF !important;
  background-color: #9F2042;
}

#main .nav-pills > li.active > a,
#main .nav-pills > li.active > a:hover,
#main .nav-pills > li.active > a:focus {
  color: #FFFFFF !important;
  background-color: #9F2042;
}

.modal-lg .modal-header {
  padding: 0;
}

.modal-lg .modal-body {
  padding: 0;
}

.modal-lg .modal-footer {
  padding: 0;
}

</style>

<!--
<script>
MathJax = {
  // your configuration here, if needed
};
</script>
-->

<script type="text/javascript" id="MathJax-script" async
  src="https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml.js">
</script>

```{r include=FALSE}

# Check and install missing libraries

list.of.packages <- c("ggplot2","dplyr", "DT", "readxl", "stringr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[
  ,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Remove temporary variables

rm(list.of.packages)
rm(new.packages)

# Manual packages (run once)
# remotes::install_github("juba/rmdformats")
```

```{r include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, echo=FALSE)
```

```{r}
library(bsplus)
library(htmltools)
bs_modal(id = "modal_markdown", title = "", size="large", 
  body = htmltools::tags$table(height="100%", tags$iframe(src="https://www.desmos.com/scientific", width="100%", height="350", scrolling="no")))
```

## Anything to the power of zero is 1.

<div class='highlight_blue'>
<strong>Any _base_ with an _exponent_ of 0 will result in 1.

Always.</strong>
</div>



## Examples

<div class='highlight_red'>

$$0^0=1$$

$$5^{0}=1$$

$$(-25)^{0}=1$$

$$385,396,394,917.95^{0}=1$$


$$(-193,395.918351202)^{0}=1$$
</div>


## Practice 

<div class='highlight_red'>

$$7^0$$

<details><summary>Show Answer</summary>

$$7^0=1$$

</details>



$$(-8)^0$$

<details><summary>Show Answer</summary>

$$(-8)^0=1$$

</details>



</div>
